Co-citations in context

### 2019-09-15 Revision2Letter.tex
tweaks

\vspace{5mm}
\raggedright 
Thank you for the opportunity to submit a revision. Again, we thank the reviewers for their diligence and professional courtesy. We have considered the comments from both reviewers carefully and provide a response below. The response is also reflected in a revised manuscript with blue text indicating where changes have been made to the first revision. We have also responded to your suggestion that we include a discussion of the work of Wang, Veugelers, and Stephan.

@@

\emph{The revised manuscript is much improved, and the specific comments to both reviews were very helpful. Overall, the authors have addressed all of the issues I’ve raised satisfactorily. The paper is important as a critique to Uzzi.The paper does not claim to present an alternative (better) way to identify innovative papers.  But I was hoping that it could- and agree with the editor that the additional requirement (of comparing evaluations from both approaches) is not needed.
In my second reading, I did realize one additional flaw that I would like the authors to address- upon the editor’s discretion.  It looks like the study had a low threshold (just two references) for including papers in their studies. And we know that citation impact is correlated with the number of references. Is it possible that the signals they (or Uzzi) get for ‘innovative’ or ‘conforming’ are dependent on the number of references? This can be easily checked (a simple table of \# of references and likelihood of a signal of conformity or innovativeness) should be done. I suspect you won’t get any signal for papers with fewer than 5 references (how can one really assess either concept with such a small number?).  But with 10 or more references- I'm quite comfortable with conclusions about impact being based on conformity and innovativeness. They may also find, when applying a reference threshold (such as 10) that they get different results.}

\begin{itemize}
\item The reference count issue is interesting although it is not quite clear what is meant by signal. We chose to set the threshold at 2 references since it is the smallest number that yields a reference pair. It is perhaps useful to be reminded that these reference pairs are then mapped to journal pairs whose frequencies are summed across the dataset being studied, then normalized, and finally returned to the parent article as z-scores. The reviewer argues that 10 references (45 z-scores) would be a reassuring number. This choice appears intuitive rather than reasoned but likely reflects insight. However, there is a penalty for arbitrarily excluding low-reference publications. For example, restricting the applied physics dataset from 1985 we analyze (ap\_1985) to those with 5 references or less results in a loss of 28\% of the publications and their associated references. \\

\vspace{1mm}
We reviewed the literature in which correlation between reference count and citations is claimed. In particular, we studied Peters and van Raan (1994) and Vieira and Gomes (2010) since other articles reporting a correlation between reference count and citations seem to eventually trail back to these two. 
\begin{enumerate}
\item Peters and van Raan used a set of 226 articles drawn from the bibliographies of 18 internationally recognized chemical engineers for the years 1977-1986. They observed a possible bi-phasic distribution when 5-year impact was measured in a histogram with reference count intervals of 5 between groups ($<=9, 10-14, 15-19, 20-24, 35-29, 30-34, >=35$). They reported a Pearson's correlation coefficient of 0.52 and inferred a weak but positive correlation (Fig 1d and Table 1 of their paper) consistent with the observations of Moed \emph{et al} 1985. This is not very convincing (or easily compared to our larger datasets from later years) but we do respect the finding.

\item More interestingly, Vieira and Gomes examined 226,166 articles from 2004 classified into Biology \& Biochemistry, Chemistry, Mathematics, and Physics per the WOS ESI. Biology and Biochemistry consisted of 455 journals, 44248 articles; Chemistry (574j, 97177a); Mathematics (387j, 20127a); and Physics (338j, 64614a). Our study differs in that we are looking at different subsets of the data, larger amounts of data, and counting citations for the first 8 years.

As far as we can tell, these authors counted total citations accumulated in around 5 years (the paper was submitted and accepted in 2009).  Fig 9 of their paper shows a positive relationship between the number of references and the citations that differs in slope and intercept between subjects when a linear fit is applied to the data.  We evaluated the relationship between references and citation counts on all 9 local networks we analyzed in this paper using the approach of Vieira and Gomes. The results are very interesting. When plotting reference counts (independent variable on the x axis) against the mean of citation counts we see a linear relationship that gets noisy after about papers with 75 references or more (our Figure 1 below). This is consistent with the observations of Vieira and Gomes and not unexpected since the noisy part of the data could be accounted for by review articles. However, when plotting citation counts (independent variable on the x axis) against the mean of reference counts the results suggest that publications with a broad range of cited references can accrue similar citations over the first 8 years since publication (our Figures 2 \& 3). We conclude that Fig 9 of Vieira and Gomes masks the variance of citation counts when the mean citation count is plotted as a dependent variable. Thus, the basis for arguing that reference counts correlate with citation counts is a little shaky although it may be true for a narrow range and subset of articles (Figure 3, column 3 in particular). In other words, trying to focus the analysis on just a subset of the data based on number of references is not likely to provide much insight. Nor will studying this be simple as not only reference counts but a number of factors would seem to explain the variance (both Peters and van Raan \& Vieira and Gomes).

However, controlling for the number of references (as a proxy for the number of z-scores) in a publication appears reasonable so we provide some data below. Perhaps this is what the reviewer had in mind as well?
\end{enumerate}
\item INSERT Jim's data
\end{itemize}

@@

\begin{itemize}
\item An illustration has been generated that will be added to supplementary material that contains material from the first rebuttal.
\item Yes, the reviewers description of our citation swapping strategy is now correct. Further, a difference between bibliometrics and mathematics may be that bibliometrics papers cite mathematics papers more often than the other way around. The local model accounts for this but the global one does not. We argue consistently that local models reduce mis-specification and provide data to support the argument.
\item The relationship between novelty, conventionality, and citation counts is not a ground truth, we have never claimed it is. This framework was developed by Uzzi et al. (2013) and we have used it to enable \emph{ceteris paribus} comparisons. We cannot address the question of the `huge assumption', which is not actually stated by the reviewer- we don't think we make such an assumption.
\end{itemize}

\subsection{Editor}
\emph{I have one other suggestion. In addition to the work on novelty by Brian Uzzi and colleagues, there is also work on novelty by Reinhilde Veugelers and colleagues, who operationalize novelty in a different way. For completeness, you may consider providing a brief discussion of the work by Veugelers and colleagues in your manuscript.}
\begin{itemize}
\item We thank you for bringing this publication to our attention. We really ought to have cited it but we were guilty of, perhaps myopically, focusing on the details of the random graph approach. Discussing this paper broadens the context of our manuscript and we've now referenced it in the Discussion. As one of our co-authors observes, the definition of novelty is a little restrictive to say the least but the paper is an excellent one and we are happy to reference it.
\item We have also added a sentence about the limitations of grouping by journals and have suggested that analyzing article level clusters may be more insightful. This has come out of our more recent interest in clustering, which informs this new direction.
 \end{itemize}

>>>>

\vspace{5mm}
\raggedright 
Thank you for the opportunity to submit a revision. Again, we thank the reviewers for their diligence and professional courtesy. We have considered their comments and provide a response below. We have also responded to your suggestion that we include a discussion of the work of Wang, Veugelers, and Stephan. Our response is reflected in a revised manuscript with blue text indicating where changes have been made to the first revision or where we have made minor edits to assist clarity.

@@

\emph{The revised manuscript is much improved, and the specific comments to both reviews were very helpful. Overall, the authors have addressed all of the issues I’ve raised satisfactorily. The paper is important as a critique to Uzzi. The paper does not claim to present an alternative (better) way to identify innovative papers.  But I was hoping that it could- and agree with the editor that the additional requirement (of comparing evaluations from both approaches) is not needed. In my second reading, I did realize one additional flaw that I would like the authors to address- upon the editor’s discretion.  It looks like the study had a low threshold (just two references) for including papers in their studies. And we know that citation impact is correlated with the number of references. Is it possible that the signals they (or Uzzi) get for ‘innovative’ or ‘conforming’ are dependent on the number of references? This can be easily checked (a simple table of \# of references and likelihood of a signal of conformity or innovativeness) should be done. I suspect you won’t get any signal for papers with fewer than 5 references (how can one really assess either concept with such a small number?).  But with 10 or more references- I'm quite comfortable with conclusions about impact being based on conformity and innovativeness. They may also find, when applying a reference threshold (such as 10) that they get different results.}

\begin{itemize}
\item We reviewed the literature in which correlation between reference count and citations is claimed. In particular, we studied Peters and van Raan (1994) and Vieira and Gomes (2010) since other articles reporting a correlation between reference count and citations seem to eventually trail back to these two. 
\begin{enumerate}
\item Peters and van Raan used a set of 226 articles drawn from the bibliographies of 18 internationally recognized chemical engineers for the years 1977-1986. They observed a possible bi-phasic distribution when 5-year impact was measured in a histogram with reference count intervals of 5 between groups ($<=9, 10-14, 15-19, 20-24, 35-29, 30-34, >=35$). They reported a Pearson's correlation coefficient of 0.52 and inferred a weak but positive correlation (Fig 1d and Table 1 of their paper) consistent with the observations of Moed \emph{et al} 1985. This is not very convincing (or easily compared to our larger datasets from later years) but is pertinent.

\item More interestingly, Vieira and Gomes examined 226,166 articles from 2004 classified into Biology \& Biochemistry, Chemistry, Mathematics, and Physics per the WOS ESI. Biology and Biochemistry consisted of 455 journals, 44248 articles; Chemistry (574j, 97177a); Mathematics (387j, 20127a); and Physics (338j, 64614a). As far as we can tell, these authors counted total citations accumulated in around 5 years (the paper was submitted and accepted in 2009).  Our study differs in that we look at different subsets of the data, larger amounts of data, and count citations for the first 8 years. \\
\vspace{3mm}
Fig 9 of their paper shows a positive relationship between the number of references and the citations that differs in slope and intercept between subjects when a linear fit is applied to the data.  We evaluated the relationship between references and citation counts on all 9 local networks we analyzed in this paper using the approach of Vieira and Gomes. The results are very interesting. When plotting reference counts (independent variable on the x axis) against the mean of citation counts we see a linear relationship that gets noisy after about papers with 75 references or more (Figure 1 of this document below). This is consistent with the observations of Vieira and Gomes and not unexpected since the noisy part of the data could be accounted for by review articles. However, when plotting citation counts (independent variable on the x axis) against the mean of reference counts the results suggest that publications with a broad range of cited references can accrue similar citations over the first 8 years since publication (our Figures 2 \& 3). We conclude that Fig 9 of Vieira and Gomes masks the variance of citation counts when the mean citation count is plotted as a dependent variable.

Thus, the basis for arguing that reference counts generally correlates with citation counts seems a little shaky although it may be true for a narrow range and subset of articles (Figure 3, column 3 in particular). A number of factors help explain variance in citation counts (both Peters and van Raan \& Vieira and Gomes) and focusing on reference count alone is distracting to the purpose of our paper.

\item Nevertheless, the reference count issue raised is interesting in itself (although it is not quite clear what is meant by signal) so we have added a sentence to the Discussion addressing it. The reviewer argues that 10 references (45 z-scores/article) would be a reassuring number. This choice appears intuitive rather than reasoned. We chose to set the threshold at 2 references since it is the smallest number that yields a reference pair. This threshold also supports comparison with Uzzi's data. It is perhaps useful to be reminded that these reference pairs are then mapped to journal pairs whose frequencies are summed across the dataset being studied, then normalized, and finally returned to the parent article as z-scores. Lastly, there is a penalty for arbitrarily excluding low-reference publications. For example, restricting the applied physics dataset from 1985 (ap\_1985) to publications with 5 references or less results in a loss of 28\% of the publications and associated references.

\vspace{3mm} Even so, \emph{we INSERT Jim's new data here... }

\begin{table}[ht]
\begin{centering}
\caption{Hit Rates By Category: Effect of Omitting Articles with 10 and Fewer References}
\vspace{3mm}
\begin{tabular}{|c crr rrr r|}
 \hline
 Dataset & Year & Hit Rate & Omitted  & LNLC & LNHC & HNLC & HNHC \\ 
  \hline
Metabolism & 2005&1\%&& 0.6& \bf{1.3}& 0.7& \bf{1.3} \\ 
Metabolism & 2005&1\%& \checkmark & 0.7& \bf{1.4}& 0.7& \bf{1.2} \\ 
\hline
Metabolism & 2005&2\%&& 1.0& \bf{2.5}& 1.5& \bf{2.6} \\ 
Metabolism & 2005&2\%& \checkmark & 1.2& \bf{2.7}& 1.5& \bf{2.5} \\ 
\hline
Metabolism & 2005&5\%&& 2.3& 6.0& 3.9& \bf{6.6} \\ 
Metabolism & 2005&5\%& \checkmark & 2.7& \bf{6.3} & 3.8& \bf{6.4} \\ 
\hline
Metabolism & 2005&10\%&& 4.1& 11.4& 8.1& \bf{12.6} \\ 
Metabolism & 2005&10\%& \checkmark & 5.3& \bf{12.0} & 8.1& \bf{12.5} \\ 
\hline
\end{tabular} \\
\end{centering}
\end{table} 
\end{enumerate}
\end{itemize}

@@

\begin{itemize}
\item An illustration has been generated that will be added to supplementary material that contains figures from the first rebuttal.
\item Yes, the reviewer's description of our citation swapping strategy is now correct. Further, a difference between bibliometrics and mathematics may be that bibliometrics papers cite mathematics papers more often than the other way around. The local model accounts for this but the global one does not. We argue consistently that local models reduce mis-specification and provide data to support the argument.
\item We agree with the reviewer that Uzzi et al. make an assumption regarding this connection between novelty, conventionality, and citation counts. We have not portrayed this assumption as a ground truth in the manuscript. This framework was developed by Uzzi et al. (2013) and we have used it to enable \emph{ceteris paribus} comparisons between our work and prior art. Furthermore, the analysis we provided using K-L divergence (between a generative model and the empirical data) establishes that there is a reduction in model misspecification by restricting to local networks and is independent of this assumption.
\end{itemize}

\subsection{Editor}
\emph{I have one other suggestion. In addition to the work on novelty by Brian Uzzi and colleagues, there is also work on novelty by Reinhilde Veugelers and colleagues, who operationalize novelty in a different way. For completeness, you may consider providing a brief discussion of the work by Veugelers and colleagues in your manuscript.}
\begin{itemize}
\item We thank you for bringing this publication to our attention. We really ought to have mentioned it but we were focused on the details of the random graph approach in local and global contexts. Discussing this paper broadens the context of our manuscript and we've now referenced it in both the Introduction and Discussion. The definition of novelty is a little restrictive in this paper but it makes a number of excellent points.
\item We added a sentence about the limitations of grouping by journals and have suggested that analyzing article level clusters may be more insightful. We have also sounded a cautionary note (that ought to be superfluous) about the limitations of using citation counts as a readout for impact. \end{itemize}

### 2019-09-15 cocit_r_v2.tex
tweaks

The principal difference between the two models we discuss is a single parameter--the set of references that can be used as substituents during the substitution process. The keyword search we use also has the advantage of selecting only relevant articles from multidisciplinary journals. However, it is important to note that the local networks we evaluated are not monodisciplinary, the references cited within exhibit disciplinary diversity. We provided several lines of evidence that showed that changing this one parameter from a global network to the local disciplinary network reduces model misspecification. Using the constrained model (which allows substitutions only within the local network)  instead of the unconstrained model (which allows substitutions in the WoS network) produces different trends in terms of conventionality and novelty, depending on the network and the parent discipline. In particular, when using the unconstrained model, highly cited papers were most likely to be in the HNHC category but this trend does not consistently hold when using the constrained model. Instead, we find that conventionality flavored with novelty is {\em not} generally a feature of impactful research. Further,  high ``novelty" is not always indicative of impactful research (as reflected in citation counts), as shown by both Immunology and Metabolism having their highest hit rates for low novelty (Figure \ref{fig:Fig2}).

@@

Our work has shown that the use of local networks enables simulations that are more consistent with research citation patterns. Further work might explore additional constraints on random assignment of citations to publications to better align benchmarks with citation practice. For example, proximity defined by co-author networks~\citep{wallace_lariviere_gingras_2012} might be considered when defining probabilities for citation substitutions. Another interesting but challenging direction would be to find ways to distinguish intra-disciplinary from cross-disciplinary novelty. \textcolor{blue}{In this respect, the related work of \citep{wang_bias_2017} is particularly insightful with its use of empirical data and the observations made on novelty, quality, as well as dispersion and kinetics of accrued citations of articles classified as novel.}

\textcolor{blue}{We note that journals are used as grouping units for articles in the three studies we discuss \citep{wang_bias_2017, uzzi_atypical_2013, boyack_vs_uzzi_2014} as well as this one. While we used keyword searches to identify sets of articles, we still relied on journal grouping to generate z-scores. Such grouping, while appealing on account of relative simplicity, obscures measurements of novel pairings at the article level. Journals are also of limited use in representing individual fields and repeating some of these studies using article clusters may be more informative~\cite{traag_louvain_2019, klavans_which_2017}}.  Overall, evaluation in context~\citep{hicks_bibliometrics:_2015} and increased consideration of the disciplinary nature of the scientific enterprise is likely to result in improved models that yield further insight.

>>>>

The principal difference between the two models we discuss is a single parameter--the set of references that can be used as substituents during the substitution process. The keyword search we use also has the advantage of selecting only relevant articles from multidisciplinary journals. However, it is important to note that the local networks we evaluated are not monodisciplinary, the references cited within exhibit disciplinary diversity. We provided several lines of evidence that showed that changing this one parameter from a global network to the local disciplinary network reduces model misspecification. Using the constrained model (which allows substitutions only within the local network)  instead of the unconstrained model (which allows substitutions in the WoS network) produces different trends in terms of conventionality and novelty, depending on the network and the originating discipline. In particular, when using the unconstrained model, highly cited papers were most likely to be in the HNHC category but this trend does not consistently hold when using the constrained model. Instead, we find that conventionality flavored with novelty is {\em not} generally a feature of impactful research. Further,  high ``novelty" is not always indicative of impactful research (as reflected in citation counts), as shown by both Immunology and Metabolism having their highest hit rates for low novelty (Figure \ref{fig:Fig2}).

@@

Our work has shown that the use of local networks enables simulations that are more consistent with research citation patterns. Further work might explore additional constraints on random assignment of citations to publications to better align benchmarks with citation practice. For example, proximity defined by co-author networks~\citep{wallace_lariviere_gingras_2012} might be considered when defining probabilities for citation substitutions. Another interesting but challenging direction would be to find ways to distinguish intra-disciplinary from cross-disciplinary novelty. \textcolor{blue}{In this respect, the related work of \citep{wang_bias_2017} is insightful with its use of empirical data and observations made on novelty and quality, as well as dispersion and kinetics of accrued citations of articles classified as novel.}

\textcolor{blue}{We note that journals are used as grouping units for articles in the three studies we discuss \citep{wang_bias_2017, uzzi_atypical_2013, boyack_vs_uzzi_2014} as well as this one. While we used keyword searches to identify sets of articles, we still relied on journal grouping to generate z-scores. Such grouping, while appealing on account of relative simplicity, obscures measurements of novel pairings at the article level. Journals are also of limited use in representing individual fields and repeating some of these studies using article clusters may be more informative~\cite{traag_louvain_2019, klavans_which_2017}. We also acknowledge the limitations of using citation counts to identify impactful publications. Various factors contribute to citations \citep{peters_determinants_1994, vieira_citations_2010} and further study in the context of co-citation may be merited. Overall, evaluation in context~\citep{hicks_bibliometrics:_2015} and further consideration of the disciplinary nature of the scientific enterprise is likely to result in improved models that yield further insight.}
